---
output: html_document
---

1. Loading and pre-processing the data
======================================

```{r readData, echo=TRUE}
fileName <- paste(getwd(),"/","repdata-data-activity","/","activity.csv",sep="")

activity <- read.csv(file = fileName, header=T)

str(activity)
ncount <- nrow(activity)
```
The no. of observations is `r ncount`.

2. What is mean total number of steps taken per day?
====================================================

```{r meanMedian, echo=TRUE}
library(sqldf)
options(scipen=999)
summaryData <- sqldf("SELECT activity.date, SUM(activity.steps) AS TOTAL_STEPS FROM activity GROUP BY activity.date")
summaryData
TOTAL_STEPS <- as.numeric(summaryData$TOTAL_STEPS)
mean <- mean(TOTAL_STEPS, na.rm=TRUE)
median <- median(TOTAL_STEPS,na.rm=TRUE)
```


Drawing a histogram depicting frequency distribution of no. of steps for the entire duration

```{r histogram, echo=TRUE}
summaryData$TOTAL_STEPS[is.na(summaryData$TOTAL_STEPS)] = 0
numStepsVsDates <- rep(summaryData$date,summaryData$TOTAL_STEPS)
hist(as.Date(numStepsVsDates),format="%d %b %y",breaks="days",xlab="Date",ylab="No. of steps",freq=TRUE)

```

Mean is `r mean` and Median is `r median`.

3. What is the average daily activity pattern?
==============================================

```{r timeseries, echo=TRUE}
avgSummaryData <- sqldf("SELECT activity.date, AVG(activity.steps) AS AVG_STEPS FROM activity GROUP BY activity.date")
avgSummaryData$AVG_STEPS <- as.numeric(avgSummaryData$AVG_STEPS)
avgSummaryData$AVG_STEPS <- round(avgSummaryData$AVG_STEPS,0)
avgSummaryData$AVG_STEPS[is.na(avgSummaryData$AVG_STEPS)] = 0
str(avgSummaryData)
head(avgSummaryData,2)
plot(avgSummaryData$date,avgSummaryData$AVG_STEPS,type="l",xlab="date",ylab="Average no. of steps per day")

maxNumStepsDay <- avgSummaryData$date[which.max(avgSummaryData$AVG_STEPS)]

```

Day, which has maximum no. of average steps,  is `r maxNumStepsDay`.


4. Inputing missing values
===========================

```{r missingValueCount, echo=TRUE}
NAMissingCount = sum(is.na(activity$steps))

```

No. of missing values in the dataset is `r NAMissingCount`.


```{r missingValueTreatment, echo=TRUE}
#i <- 1
#for(i in 1:nrow(activity)){
 # missingValue <- activity$date[i]
 # activity$steps[is.na(activity$steps)] =     
   # avgSummaryData$AVG_STEPS[avgSummaryData$date==missingValue]
#print(missingValue)
#i <- i+1
#}

activity$steps2 <- activity$steps
ind <- is.na(activity$steps2) #find where missing values are
missingDates <- activity$date[ind] #find to what dates the missing values belong to

activity$steps2[ind] <- avgSummaryData$AVG_STEPS[missingDates] #substitute the missing values with average values of no. of steps for that date


```

Changed the missing values to average no. of steps for that day

```{r histogramTreatedData, echo=TRUE}

hisogramData <- sqldf("SELECT activity.date, SUM(activity.steps2) AS TOTAL_STEPS FROM activity GROUP BY activity.date")

histStepsVsDates <- rep(hisogramData$date,hisogramData$TOTAL_STEPS)
hist(as.Date(histStepsVsDates),format="%d %b %y",breaks="days",xlab="Date",ylab="No. of steps",freq=TRUE)

TOTAL_STEPS <- as.numeric(hisogramData$TOTAL_STEPS)
mean <- mean(TOTAL_STEPS, na.rm=TRUE)
median <- median(TOTAL_STEPS,na.rm=TRUE)
```

New values for Mean is `r mean` and Median is `r median`.
Both mean and median values have reduced after this treatment.


5.Are there differences in activity patterns between weekdays and weekends?
============================================================================

```{r weekendWeekday, echo=TRUE}
#dataWeekend <- 
```